---
tables: yes
output: 
  beamer_presentation:
    keep_tex: yes
    theme: Antibes
    colortheme: "dolphin"
    fonttheme: "professionalfonts"
header-includes:
  - \useoutertheme{infolines}
  - \useinnertheme{circles}
  - \setbeamercolor{note page}{bg=white,fg=black}
  - \setbeamercolor{note title}{bg=white!80!black, fg=black}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#- %\setbeamercolor{note date}{parent=note title}
```

```{r eval=FALSE, include=FALSE}
output: 
  beamer_presentation:
    keep_tex: yes
    theme: Antibes
    colortheme: "dolphin"
    fonttheme: "professionalfonts"

    
      - \usetheme{metropolis}

```


# Pareto Distribution

## Background

- Vilfredo Pareto (1897) presents a versatile functional relation that well describes wealth distribution across countries and centuries. 
- Same concept is applied to several other fields and colloquially called
*Pareto Principle*.
  - 80% of land owned by 20% of individuals (revenue $\sim$ products; sales $\sim$ clients; etc)
- Generally, it follows a *power law probability distribution*, where one measure varies constantly as an exponential of another, independently of initial values.
  - Example: if one increases the side length of a square by $x$, its area increases by $x^2$, independently of initial area of square.

\note{\begin{itemize}\tightlist
\item
I am going to talk about the theoretical framework in our replication study. Since
the focus of the study is not only to provide new information, which is usually 
hard to come by, but also to check if we can make some more general inferences 
about  wealth distribution in germany and adapt the top tail of the SOEP accordingly. 
\item AS A BACKGROUND: Pareto (the same of the optimallity concept)
\item functional form that describes well the wealth distribution accross countries and time

\item an approximate description of the distribution of incomes and wealth among the rich and the moderately rich
\item 80/20 rule (revenue / sales; )
\item Latent sector errors in hard-drives failures and Clusters of Bose–Einstein condensate near absolute zero

\end{itemize}}

## Functional Form

The **Pareto Distribution** is defined by 
$$f(y, \underline y, \alpha) = \frac{\alpha \underline y^\alpha}
{y^{\alpha+1}} , \quad 0 < \underline y < y$$

and
$$F(y, \underline y, \alpha ) = 1 - \left( \frac{~\underline y~}{y}   \right)^{\alpha} , \quad 0 < \underline y < y$$

where:

- $y$: wealth measure
- $\underline y$: lower bound (or *scale parameter* or *threshhold value*)
- $\alpha$: Pareto's $\alpha$ (or *shape*/*slope parameter*)


\note{\begin{itemize}\tightlist
\item important here is only that
  \item - only defined for postive values
  \item - and above the lower-bound
  \item - given a certain lower-bound, alpha is the important paramenter, known as
  pareto alpha
.\\ 
.\\
.\\
The lower-bound can raise some issues, as we are going to see in the next slides
\end{itemize}}


## Graphical Visualisation I

![](./graphs/04_paretoDistGraphs.pdf){height="300px"}

\note{\begin{itemize}\tightlist

\item Here we can see how the Probability and Cumulative Distribution looks like.
\item - Only defined after lowerbound
\item - very heavy right tail 
\item - very left dense

\end{itemize}}


## Graphical Visualisation II{.fragile}

![](graphs/04_paretoDiagram.pdf){height="300px"}

\note{\ttfamily{
In order to better visualise this distribution TWO transformations are taken:

1. calculate 1 - CDF\\
2. plot on log-log scale (or log-log the data)\\
\qquad .\\
\qquad .

HERE: lower bound = 1000 \\
\qquad $~~\alpha$ = Slope in loglog \\
\qquad - Approx. interpretation: for a percentage increase in $y$, the proportion of *richer* individuals by $\alpha$ percents (NOT PERCENTAGE POINTS).
}}


## Properties 

*Pareto's $\alpha$:*

- Approx. interpretation: for a percentage increase in $y$, the proportion of *richer* individuals by $\alpha$ percents.
- Higher $\alpha$ values $\Rightarrow$ less inequality.^[for inequality measures satisfying the *Weak Transfers Principle* (Cowell 2011, p. 93).]
- Several inequality indices can be estimated based on $\alpha$.
  - Example: Gini coefficient: $\frac{1}{2 \alpha - 1}$.

*Possible problems:*

- High flexibility on estimating the lower bound
- Sensibility of $\alpha$ due to choice of the lower bound $\underline y$


\note{
PROPERTIES
\begin{itemize}\tightlist
\item Interpretation
\item higher value --> less inequality
\item Inequality indices can be estimated based on $\alpha$
\item - - - GINI coef.

\end{itemize}

PROBLEMS
\begin{itemize}\tightlist
\item High flexibility on estimating the lower bound
\item Sensibility of $\alpha$ due to choice of the lower bound $\underline y$
\item For narrow excerpts of the data other distributions are "just as good"
\end{itemize}
}

