---
tables: yes
output: 
  beamer_presentation:
    keep_tex: yes
    theme: Antibes
    colortheme: "dolphin"
    fonttheme: "professionalfonts"
    slide_level: 3
header-includes:
  - \usepackage{adjustbox}
  - \useoutertheme{infolines}
  - \useinnertheme{circles}
  - \setbeamercolor{note page}{bg=white,fg=black}
  - \setbeamercolor{note title}{bg=white!80!black, fg=black}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#- %\setbeamercolor{note date}{parent=note title}
```

```{r eval=FALSE, include=FALSE}
output: 
  beamer_presentation:
    keep_tex: yes
    theme: Antibes
    colortheme: "dolphin"
    fonttheme: "professionalfonts"

    
      - \usetheme{metropolis}
    
    
    
"tables/alhpa_results_p95.tex"
```

# DO NOT INLCUDE THIS 

TODO: dont include it

# INCLUDE THIS

## Hausman Test

### Robustness Check of Pareto's alpha {.allowframebreaks}



Question: Is the Pareto distribution a *good enough* approximation of the 
wealth distribution among Germany's top wealth individuals?

If this is the case, one can expect that: 

- the $\alpha$ parameter is across SOEP and Pretest datasets and
- among the chosen lower bounds, while taking the uncertainty from the
imputation method into account.

To check this hypothesis we perfom a generalized Hausman Test based on Hausman 
(1978) and Hausman and McFadden (1984).


### Generalized Hausman Test

We test the stability of the estimated Pareto's $\alpha$ by cross checking the
obtained coefficients among a combination of each dataset, percentile and
imputed data, via stata's `suest` post-estimation command.


$$H_0: \hat\alpha^{data_i}_{pct_k, imp_m} = \hat\alpha^{data_j}_{pct_l, imp_m}$$

where $\textit{data}~in~\{SOEP, Pretest\} \text{ datasets}$, 
$\textit{pct}~in~\{95th, 99th\} \text{ percentiles}$ and 
$\textit{imp}~in~\{1, ... 5\} \text{ imputations}$.


The Hausman test statistic follows

$$ \frac{  \left( \hat\alpha_i - \hat\alpha_j \right)^2 }
    {var(\hat\alpha_i) - 2  cov(\hat\alpha_i,\hat\alpha_j) 
      + var(\hat\alpha_j)} \sim \chi^2_{1}.
$$

\note{\begin{itemize}\tightlist

\item we use Hausman via suest because we are able to compare coefficients 
of two distinct regressions and grouping the data would be problematic due to
undefined reweighting scheme for combined dataset

\end{itemize}}




## Test Results 

### Results I 


\begin{table}[]
  \centering
  \resizebox{\textwidth}{!}{\input{tables/alhpa_results_p95.tex}}
  \caption{Estimation of Pareto's $\alpha$ across data and imputated sets
  for the lower bound at the 95th percentile.}
    \label{tab:my_label}
\end{table}

### Results II


\begin{table}[]
        \centering
        \resizebox{\textwidth}{!}{\input{tables/alhpa_results_p99.tex}}
        \caption{Estimation of Pareto's $\alpha$ across data and imputated sets for the 
        lower bound at the 99th percentile.}
        \label{tab:my_label}
\end{table}




