---
title: "Re-Weighting Stratified Sampling Example"
author: "Marcelo Avila"
date: "29 November 2018"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(EnvStats) # to generate random pareto data
```


## Generate fake data following pareto distribution 

```{r}
set.seed(42)
shape = 2   # shape = alpha param
location = 220 # location = lower bound

popAll <- EnvStats::rpareto(12500,  location = location, shape = shape)
(summary(popAll))
```

### define strata break points 

```{r}
b1 <- qpareto(.33, location = location, shape = shape) # lower third
b2 <- qpareto(.66, location = location, shape = shape) # middle third

# divide into 3 strata
p1 <- popAll[popAll<=b1]
p2 <- popAll[popAll>b1 & popAll<=b2]
p3 <- popAll[popAll>b2]
```


### plot histogram 

```{r, warning=FALSE}
# plot
hist(popAll, col=rgb(1,0,0,0.5), freq = T, breaks = 500, xlim = c(location, 2000),
     main = "Histogram", xlab = "net wealth")

# plot strata bounds in red
abline(v = b1, lty=1, lwd=3, col = "red")
abline(v = b2, lty=1, lwd=3, col = "red")

# define means
mean1 <- mean(p1)
mean2 <- mean(p2)
mean3 <- mean(p3)

# plot means in blue
abline(v=mean1, col="blue", lty=2)
abline(v=mean2, col="blue", lty=2)
abline(v=mean3, col="blue", lty=2)
```


## Stratification Method

```{r}
# define weights for disproportional Strata
gew1 <- 1/7
gew2 <- 2 * gew1
gew3 <- 4 * gew1

ss <- 2100     # total sample size
s1 <- sample(p1, size = gew1 * ss)
s2 <- sample(p2, size = gew2 * ss)
s3 <- sample(p3, size = gew3 * ss)

c(n1=length(s1), n2=length(s2), n3=length(s3)) ## size of each sampling
```


### Plot histogram of Samples

```{r}
breaks = 1:100000 * 20
hist(s1, col=rgb(1,0,0,0.5), main = "histogram of Samples",
     xlim = c(location, 2000), ylim = c(0,150), breaks = breaks,
     xlab = "red: s1, green: s2, blue: s3") 
     
hist(s2, col=rgb(0,1,0,0.5), breaks = breaks, add=T)
hist(s3, col=rgb(0,0,1,0.5), breaks = breaks, add=T)
```

**note:** the overlapping is rather an artifact of the histogram bins construction, 
since the data itself do not overlap.

```{r}
# strata means
c(s1mean = mean(s1), s2mean = mean(s2), s3mean= mean(s3))
```

```{r}
# population mean
c(mean(p1), mean(p2), mean(p3))
```


### Stratified Mean Estimation 

```{r}
# sample and pop sizes
N1 <- length(p1)
N2 <- length(p2)
N3 <- length(p3)
N <- sum(N1, N2, N3)

# unbiased estimation 
strataMean <- (1/N) * sum( c(N1,N2,N3 ) * c( mean(s1), mean(s2), mean(s3) ))

```

with strataMean being calculated as $\bar x = \frac{1}{N} \sum_h^L N_h \bar x_h$

where:

- $L$: count of strata, 
- $N$: number of population observations,
- $N_h$: number of observations in stratum $h$,
- $\bar x_h$: mean of stratum $h$,
- $h$ = 3 in our case.

**Interesting to see:** its not the inverse of *gew* (1/7, 2/7 and 4/7) that
gets to rewheight the averages of the strata mean into the overall mean, but the
proportion of of observations in each strata compared to the whole poplation (n1
/ N) and they are all around 1/3 and not related to the specific wheights


```{r}
c(popMean = mean(popAll),
  strataMean = strataMean,
  diff = mean(popAll) - strataMean,
  percentDiff = (mean(popAll) - strataMean)/mean(popAll))
```

I suppose the rather big difference is due to the high volatility of maximum
value of the third strata. (changing the seed gives substantially different
estimates)

```{r}
# alternative calc
wgt1 <- N1 / N
wgt2 <- N2 / N
wgt3 <- N3 / N
strataMean2 <- sum(wgt1 * mean(s1), wgt2 * mean(s2), wgt3 * mean(s3))
strataMean2

# naive mean
mean(c(s1, s2, s3))
```


